---
title: "imputeTS"
author: "Dra. Martha Elva Ramírez Guzmán"
date: "2024-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<!-- # na_mean: Replaces each missing value with the mean of the present values -->
<!-- # na_locf: Replaces each missing value with the most recent present value prior to it -->
<!-- # na_kalman: Replaces each missing value with a Kalman smoothing -->
<!-- # na_seadec: Replaces each missing value with a seasonal decomposition. -->
<!-- #            The seasonal decomposition is done by stl() from the stats package. -->
<!-- #            Removes the seasonal component from the time series,  -->
<!-- #            performs imputation on the deseasonalized -->
<!-- #            series and afterwards adds the seasonal component again. -->
<!-- # na_ma: Replaces each missing value with the moving average -->
<!-- # na_random: Replaces each missing value with a random value -->
<!-- # na_remove: Removes all missing values -->
<!-- # na_replace: Replaces each missing value with a specified value -->
<!-- # na_seasplit: Splits the times series into seasons and afterwards  -->
<!-- #              performs imputation separately for each of the -->
<!-- #              resulting time series datasets (each containing the data for one specific season) -->
<!-- #              the data for one specific season). -->
```{r, message=FALSE, warning=FALSE}
# Load libraries
library(imputeTS)
library(ggplot2)
library(dplyr)
```
Functions: ggplot_na_imputations for the next types of imputations:

na_mean: Replaces each missing value with the mean of the present values.

na_locf: Replaces each missing value with the most recent present value prior to it.

na_kalman: Replaces each missing value with a Kalman smoothing.
na_seadec: Replaces each missing value with a seasonal decomposition.
           The seasonal decomposition is done by stl() from the stats package.
           Removes the seasonal component from the time series,
           performs imputation on the deseasonalized
           series and afterwards adds the seasonal component again.
           
na_ma: Replaces each missing value with the moving average.

na_random: Replaces each missing value with a random value.

na_remove: Removes all missing values.

na_replace: Replaces each missing value with a specified value.

na_seasplit: Splits the times series into seasons and afterwards
             performs imputation separately for each of the
             resulting time series datasets (each containing the data for one specific season) the data for one specific season).
# 1.1 Function: ggplot_na_distribution
```{r}
x <- stats::ts(c(1:11, 4:9, NA, NA, NA, 11:15, 7:15, 15:6, NA, NA, 2:5, 3:7))
ggplot_na_distribution(x)
```

# 1.2 Visualize NAs in tsAirgap 

```{r}
ggplot_na_distribution(tsAirgap, subtitle = NULL) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
```
  

# 2. Function: ggplot_na_ditribution2. For large time series

```{r}
ggplot_na_distribution2(tsAirgap) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::theme(plot.subtitle = ggtext::element_markdown(hjust = 0.5))
```

# 3. Function: ggplot_na_gapsize
```{r}
ggplot_na_gapsize(tsNH4, color_total = "grey")
```


# 4. Functions: ggplot_na_imputations for the next types of imputations: 
# 4.1 na_mean

```{r}
imp_mean <- na_mean(tsAirgap)
ggplot_na_imputations(tsAirgap, imp_mean)
```

```{r}
imp_mean <- na_mean(tsAirgap)
ggplot_na_imputations(x_with_na = tsAirgap,
                      x_with_imputations = imp_mean,
                      subtitle = NULL) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
```

# 4.2 na_locf and added ground truth
```{r}
imp_locf <- na_locf(tsAirgap)
ggplot_na_imputations(x_with_na = tsAirgap,
                      x_with_imputations = imp_locf,
                      x_with_truth = tsAirgapComplete
)

imp_locf <- na_locf(tsAirgap)
ggplot_na_imputations(x_with_na = tsAirgap, x_with_imputations = imp_locf)

```

# 4.3 na_kalman
```{r}
imp_kalman <- na_kalman(tsAirgap)
ggplot_na_imputations(x_with_na = tsAirgap, x_with_imputations = imp_kalman)
```

# 4.4 na_seadec
```{r}
imp_seadec <- na_seadec(tsAirgap) 
ggplot_na_imputations(x_with_na = tsAirgap, x_with_imputations = imp_seadec) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::theme(plot.subtitle = ggplot2::element_text(hjust = 0.5))
```

# 4.5.1 na_ma: Perform imputation with simple moving average
```{r}
imp_na_ma<-na_ma(tsAirgap, weighting = "simple")
ggplot_na_imputations(x_with_na = tsAirgap,
                      x_with_imputations = imp_na_ma,
                      subtitle = NULL) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
```

# 4.5.2: Imputation with EXPONENCIAL weighted moving average window size=6
```{r}
imp_na_ma_6<-na_ma(tsAirgap, k = 6)
ggplot_na_imputations(x_with_na = tsAirgap,
                      x_with_imputations = imp_na_ma_6,
                      subtitle = NULL) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
```

# 4.5.3: Same as example 1, just written with pipe operator
```{r}
tsAirgap %>% na_ma(weighting = "simple") %>% 
  ggplot_na_imputations(x_with_na = tsAirgap,
                        x_with_imputations = imp_na_ma,
                        subtitle = NULL) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
```

# 4.6 na_random
```{r}
imp_random <- na_random(tsAirgap)
ggplot_na_imputations(x_with_na = tsAirgap, x_with_imputations = imp_random)
```

# 4.7.1 na_replace. Replace all NAs with 0
# Prerequisite: Create Time series with missing values
```{r}
x <- ts(c(2, 3, NA, 5, 6, NA, 7, 8))
imp_replace<-na_replace(x, fill = 0)
ggplot_na_imputations(x_with_na = x, x_with_imputations = imp_replace)

x %>% na_replace(fill = 0) %>%
  ggplot_na_imputations(x_with_na = x, x_with_imputations = imp_replace)
```

# 4.7.2: Same as example 1, just written with pipe operator
```{r}
x <- ts(c(2, 3, NA, 5, 6, NA, 7, 8))
x %>% na_replace(fill = 3.5) %>%
  ggplot_na_imputations(x_with_na = x, x_with_imputations = imp_replace)
```

# 4.8.1 Perform seasonal splitted imputation using algorithm = "interpolation"
```{r}
imput_splitted<-na_seasplit(tsAirgap, algorithm = "interpolation") 
tsAirgap %>%   ggplot_na_imputations(x_with_na = tsAirgap,
                        x_with_imputations = imput_splitted,
                        subtitle = NULL) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
```
